### 2021-06-30 doc/temp-sentiment-summary.tex

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{breakcites}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{textgreek}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{lineno}
\usepackage{xcolor}
% \usepackage[
% style=numeric,
% citestyle=numeric-comp,
% backend=biber,
% doi=true,
% natbib=true,
% sorting=none
% ]{biblatex}
\usepackage{scalerel,xparse}

\title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}

\author[1,*]{Matthew Cooper}
\author[2]{Jeremiah Osborne-Gowey}
\author[3]{Zheng Liu}
\author[4]{Portia Adade Williams}
\author[5]{Jie Liu}
\author[6]{Aaron Schwartz}
\author[7]{Patrick Baylis}

\affil[1]{T.H. Chan School of Public Health, Harvard University}
\affil[2]{Environmental Studies Program, University of Colorado Boulder}
\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
\affil[4]{University of Cape Town}
\affil[5]{School of Business, East China University of Science and Technology}
\affil[6]{University of Colorado Boulder}
\affil[7]{University of British Columbia}
\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}

\begin{abstract}
%Summary paragraph word limit "ideally ideally of no more than 200 words" unless the editor finds the summary is compellingly enhanced by the inclusion of the broader perspective concluding sentences: https://www.nature.com/nature/for-authors/formatting-guide

% This summary paragraph should be structured as follows: 2-3 sentences of basic-level introduction to the field; a brief account of the background and rationale of the work; a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent); and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards.

% See summary paragraph example from NCC: https://www.nature.com/documents/nature-summary-paragraph.pdf

Previous research indicate strong associations between heat and human mental health \cite{Berry2018Apr, hayes_climate_2018}. Mental health represents a major share of global human suffering, accounting for up to 13\% of the total global burden of disease \cite{Collins2011Jul}. Higher temperatures associated with climate change are expected to have major impacts on mental health with researchers and medical experts sounding a clarion call for more research into potential linkages between climate change and mental health outcomes \cite{Collins2011Jul, Cardona2012Jan, Berry2018Apr}. Researchers have started responding to these calls finding associations between various environmental conditions a variety of mental health outcomes \cite{Waite2017Dec, Raker2019Dec, Cunsolo2018Apr, baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}. Yet there remains a relative paucity of empirical studies in this space. What studies exist typically use relatively coarse-scale data (e.g., city or county level) and, to date, have found no differential vulnerability for changing levels of income or race. Here, and in stark contrast to previous studies, we use finer-scale census block and climate data linked with spontaneous, in-situ observations of mood as expressed in a quarter of a billion geo-located posts on Twitter from 2009-2018, and find strong heterogeneity between heat exposure and mood with the poorest and highest proportion Black neighborhoods disproportionately negatively affected by higher temperatures. These results demonstrate some of the first and strongest to date linkages between increasing temperatures and deteriorating mental health vulnerability, as expressed through spontaneous posts about mood, and the first study that demonstrates heterogeneous effects across various income and racial and ethnic categories, particularly at finer spatial scales than previous reported. We anticipate these findings will help address the needs and risks of the most vulnerable groups and align these with research policy agendas and mental health treatment programs.
\end{abstract}

\begin{document}

\end{document}